---
title: "peer assesment 1"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Code for reading in the dataset and/or processing the data
```{r first chunk}
#load data
data <- read.csv('activity.csv')
#complete cases
activity <- data[complete.cases(data), ]
head(activity)
```
2.Histogram of the total number of steps taken each day
```{r second chunk}
#calculate steps by day
stepsbyday <- aggregate(steps ~ date, activity, sum)
stepsbyday
```
```{r histogram1, echo=TRUE}
# Histogram of steps per day
hist(stepsbyday$steps, main = "Total number of steps per day", xlab = "Steps per day")
```
3.Mean and median number of steps taken each day
```{r third chunk}
#calculate mean and median steps by day
mean(stepsbyday$steps)
median(stepsbyday$steps)
```
4. Time series plot of the average number of steps taken
```{r fourth chunk, echo=TRUE }
#calculate mean step by interval
meanstepsinterval <- aggregate(steps ~ interval, activity, mean)
```

```{r time series 1, echo=TRUE}
#Time series plot if the average number of steps taken
plot(meanstepsinterval$interval, meanstepsinterval$steps, type='l', main="Time series plot of the average number of steps taken", xlab="Interval", ylab="Mean number of steps")

```

5.The 5-minute interval that, on average, contains the maximum number of steps
```{r ad chunk}
maximunsteps<-which.max(meanstepsinterval$steps)
maximuninterval<-meanstepsinterval[maximunsteps, ]
maximunsteps
maximuninterval

```
6.Code to describe and show a strategy for imputing missing data
```{r fith chunk}
#1. Calculate and report the total number of missing values in the dataset
Missing <- length(which(is.na(activity$steps)))


#2. Devise a strategy for filling in all of the missing values in the dataset.

#3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
Imputed <- activity
for (i in 1:nrow(Imputed)) {
    if(is.na(Imputed$steps[i])) {
        val <- meanstepsinterval$steps[which(meanstepsinterval$interval == Imputed$interval[i])]
        Imputed$steps[i] <- val 
    }
}

#calculate steps by day
stepsbydayI <- aggregate(steps ~ date, Imputed, sum)
```

7.Histogram of the total number of steps taken each day after missing values are imputed
```{r histogram2, echo=TRUE}
# Histogram of steps per day
hist(stepsbydayI$steps, main = "Total number of steps per day (imputed)", xlab = "Steps per day")
```
```{r chunk bis, echo=TRUE}

mean(stepsbydayI$steps)
median(stepsbydayI$steps)

```

8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r sixth chunk}
#Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.


Imputed$date <-  ifelse(as.POSIXlt(Imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')

meanImputed <- aggregate(steps ~ interval + date, data=Imputed, mean)

```

```{r time series 2, echo=TRUE}

#Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.


library(ggplot2)
ggplot (meanImputed, aes(interval, steps)) + geom_line() + xlab("5 minute interval") + ylab("average number of steps all weekday days or weekend days ")+ labs(title= "Time series plot by weekday days or weekend days") + facet_grid(date ~ .)

    
```